### 2016-10-30 ode_two/paper.tex
working on rossler system

R\"{o}ssler's analysis focused on the behavior of the system for small $a$ and $b$. For simplicity, we can consider the solutions to (\ref{eq:csln1}) and (\ref{eq:csln2}) for very small $a$ and $b$. We now have
\begin{gather}
	\left[
	x=0,
	y=0,
	z=0
	\right]\label{eq:csln1sm}\\
	\left[
	x=c,
	y=-\frac{c}{a},
	z=\frac{c}{a}
	\right]\label{eq:csln2sm}.
\end{gather}
To find the behavior around the critical points given by (\ref{eq:csln1sm}), we evaluate the Jacobian at the appropriate points and find the eigenvalues of the system.
\[
	|J(0,0,0)-\lambda I| =
	\left|
	\left[
	\begin{array}{c c c}
		-\lambda & -1       & -1         \\
		1        & -\lambda & 0          \\
		0        & 0        & -c-\lambda
	\end{array}
	\right]
	\right| =
	(c-\lambda)(\lambda-1)(\lambda + 1)=0.
\]
Hence, the system has three eigenvalues: $\lambda=1, \lambda=-1,$ and $\lambda = c$ (we will assume here that $c > 1$). This implies that the system has an unstable saddle at the origin \cite{equilibria}. Figure \ref{fig:plot_cuts} shows various cuts of the system for various $z$ values; the system spirals in the $x-y$ plane, while moving towards and then away from the origin as $z$ moves from $-\infty$ to $\infty$. This represents only one of the equilibrium points of the system, however. We now need to analyze the linearized system near (\ref{eq:csln2sm}).
\[
	\left|J\left(c,-\frac{c}{a},\frac{c}{a}\right) -\lambda I\right| =
	\left|
	\left[
	\begin{array}{c c c}
		-\lambda    & -1          & -1       \\
		1           & a - \lambda & 0        \\
		\frac{c}{a} & 0           & -\lambda
	\end{array}
	\right]
	\right| =
	\frac{ ac - a\lambda - c\lambda + a^{2}\lambda^{2} -a\lambda^{3}}{a}.
\]
Setting the determinant to $0$ and solving in Mathematica yields three distinct (and incredibly verbose) eigenvalues with $\lambda_{1}\in\mathbb{R}$ the other two complex conjugates of each other. Such behavior lends itself towards what is known as a \textit{saddle focus}, which is always unstable \cite{equilibria}. Figure \ref{fig:3dsys_02} displays the behavior of both equilibrium points. The spiral behavior can be seen in the $x-y$ plane, and the saddle behavior can be seen predominantly in the $z-x$ plane.

@@

>>>>

R\"{o}ssler's analysis focused on the behavior of the system for small $a$ and $b$. For simplicity, we can consider the solutions to (\ref{eq:csln1}) and (\ref{eq:csln2}) for very small $a$ and $b$. We now have
\begin{gather}
	\left[
	x=0,
	y=0,
	z=0
	\right]\label{eq:csln1sm}\\
	\left[
	x=c,
	y=-\frac{c}{a},
	z=\frac{c}{a}
	\right]\label{eq:csln2sm}.
\end{gather}
To find the behavior around the critical points given by (\ref{eq:csln1sm}), we evaluate the Jacobian at the appropriate points and find the eigenvalues of the system.
\[
	|J(0,0,0)-\lambda I| =
	\left|
	\left[
	\begin{array}{c c c}
		-\lambda & -1       & -1         \\
		1        & -\lambda & 0          \\
		0        & 0        & -c-\lambda
	\end{array}
	\right]
	\right| =
	(c-\lambda)(\lambda + 1)^{2}=0.
\]
Hence, the system has three eigenvalues: $\lambda=-i, \lambda=-i,$ and $\lambda = -c$ (we will assume here that $c > 1$). This implies that the system has an unstable focus node at the origin \cite{equilibria}. Figure \ref{fig:plot_cuts} shows various cuts of the system for various $z$ values; the system spirals in the $x-y$ plane, while moving towards and then away from the origin as $z$ moves from $-\infty$ to $\infty$. We can solve this linearized system by looking at the equation
\[
	\mathbf{\dot{x}} =
	\left[
	\begin{array}{c c c}
		0 & -1 & -1 \\
		1 & 0  & 0  \\
		0 & 0  & -c
	\end{array}
	\right]
	\mathbf{x}.
\]
Since we've already found the eigenvalues of the system, we can find the eigenvectors immediately.

\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		\text{\underline{$\lambda=-i$}}\\
		\mathbf{\lambda_{1}}=\left[
		\begin{array}{c}
			-i \\
			1  \\
			0
		\end{array}
		\right].
	\end{gather*}
\end{minipage}
\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		\text{\underline{$\lambda=i$}}\\
		\mathbf{\lambda_{2}}=\left[
		\begin{array}{c}
			i \\
			1 \\
			0
		\end{array}
		\right].
	\end{gather*}
\end{minipage}
\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		\text{\underline{$\lambda=-c$}}\\
		\mathbf{\lambda_{3}}=\left[
		\begin{array}{c}
			\frac{c}{1+c^{2}}  \\
			-\frac{1}{1+c^{2}} \\
			1
		\end{array}
		\right].
	\end{gather*}
\end{minipage}

\noindent\\
\noindent Letting $\mathbf{\gamma}$ stand in for our constants, we get
\[
	\mathbf{x}(t) =
	\gamma_{1}e^{-it}\left[
	\begin{array}{c}
		-i \\
		1  \\
		0
	\end{array}
	\right] +
	\gamma_{2}e^{it}
	\left[
	\begin{array}{c}
		i \\
		1 \\
		0
	\end{array}
	\right] +
	\gamma_{3}e^{-ct}\left[
	\begin{array}{c}
		\frac{c}{1+c^{2}}  \\
		-\frac{1}{1+c^{2}} \\
		1
	\end{array}
	\right].
\]
Note that we choose to leave the $e^{it}$ in this form, since \textbf{${\lambda_{1}}$} and $\mathbf{\lambda_{2}}$ are linearly independent of each other. Figure \ref{fig:zero_sln} shows the behavior of this linearized system with two different sets of initial conditions. If we compare this to figure \ref{fig:3dsys_02}, we see similar behavior in the area of the graph with large $z$ values.

This represents only one of the equilibrium points of the system, however. We now need to analyze the linearized system near (\ref{eq:csln2sm}). In this case, we cannot let $a$ tend towards zero as we did in (\ref{eq:csln1sm}) since our roots rely on division by $a$.
\[
	\left|J\left(c,-\frac{c}{a},\frac{c}{a}\right) -\lambda I\right| =
	\left|
	\left[
	\begin{array}{c c c}
		-\lambda    & -1          & -1       \\
		1           & a - \lambda & 0        \\
		\frac{c}{a} & 0           & -\lambda
	\end{array}
	\right]
	\right| =
	\frac{ ac - a\lambda - c\lambda + a^{2}\lambda^{2} -a\lambda^{3}}{a}.
\]
Setting the determinant to $0$ and solving in Mathematica yields three distinct (and incredibly verbose) eigenvalues with $\lambda_{1}\in\mathbb{R}$ the other two complex conjugates of each other. Such behavior lends itself towards what is known as a \textit{saddle focus}, which is always unstable \cite{equilibria}. Figure \ref{fig:3dsys_02} displays the behavior of both equilibrium points. The spiral behavior can be seen in the $x-y$ plane, and the saddle behavior can be seen predominantly in the $z-x$ plane.

Again, we can solve the linearized system and analyze the behavior as we did with the system centered around $(0,0,0)$. Due to the complexity of the eigenvalues of this system, we let Mathematica determine the behavior. The command
\begin{center}
	\texttt{Eigenvectors[jcbn[x, y, z] /. \{x -> c, y -> -(c/a), z -> (c/a)\}]}
\end{center}
yields an unfriendly expression involving the \texttt{Root} command; in this case we will want to give concrete values to $c$ and $a$. Figure \ref{fig:3dsys_02} involves the behavior of the system at $a=0.14$ and $c=8.8$, so we will examine the resulting eigenvalues at those points.
\begin{center}
	\texttt{Eigenvectors[jcbn[x, y, z] /. \{x -> c, y -> -(c/a), z -> (c/a)\}]/.\{a -> 0.14, c -> 8.8\}}
\end{center}
results in the following three eigenvectors:

\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		% \text{\underline{$\lambda=-i$}}\\
		\mathbf{\lambda_{1}}=\left[
		\begin{array}{c}
			-0.0022 \\
			-1.0003 \\
			1.0000
		\end{array}
		\right]
	\end{gather*}
\end{minipage}
\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		% \text{\underline{$\lambda=i$}}\\
		\mathbf{\lambda_{2}}=\left[
		\begin{array}{c}
			0.0000174 - 0.12713i \\
			0.0159 + 0.00028i    \\
			1
		\end{array}
		\right]
	\end{gather*}
\end{minipage}
\begin{minipage}{0.3\textwidth}
	\begin{gather*}
		% \text{\underline{$\lambda=-c$}}\\
		\mathbf{\lambda_{3}}=\left[
		\begin{array}{c}
			0.0000174 - 0.12713i \\
			0.0159 + 0.00028i    \\
			1
		\end{array}
		\right].
	\end{gather*}
\end{minipage}

\noindent\\
\noindent We now have enough to for a general solution with $\mathbf{\gamma}$ in place of our constants.
\begin{gather*}
	\mathbf{x}(t) =
	\gamma_{1}e^{0.138t}\left[
	\begin{array}{c}
		-0.0022 \\
		-1.0003 \\
		1
	\end{array}
	\right] +
	\gamma_{2}e^{(0.00109-7.99i)t}\left[
	\begin{array}{c}
		0.0000174 - 0.12713i \\
		0.0159 + 0.00028i    \\
		1
	\end{array}
	\right] +\\
	\gamma_{3}e^{(0.00109+7.99i)t}\left[
	\begin{array}{c}
		0.0000174 + 0.12713i \\
		0.0159 - 0.00028i    \\
		1
	\end{array}
	\right].
\end{gather*}
Figure \ref{fig:nzero_sln} displays two parametric plots of the solution curves for different initial values. The curves exhibit a center-like behavior that looks like a spring. We can compare this to the actual curve seen in \ref{fig:3dsys_02}. This behavior is seen in the purple spiral in the $x-y$ plane of the graph.

@@

\begin{figure}[h]
	\centering
	\begin{tabular}{c c}
		\includegraphics[scale=0.5]{gen_sln_01_01} & \includegraphics[scale=0.5]{gen_sln_01_02} \\
		$\gamma_{1}=-0.7$                          & $\gamma_{1}=-0.7$                          \\
		$\gamma_{2}=-0.7$                          & $\gamma_{2}=-0.7$                          \\
		$\gamma_{3}=-1$                            & $\gamma_{3}=1$                             \\
	\end{tabular}
	\caption{Two parametric plots for the linearized system around $(0,0,0)$ with different initial values}
	\label{fig:zero_sln}
\end{figure}

\begin{figure}[h]
	\centering
	\begin{tabular}{l r}
		\includegraphics[scale=0.5]{gen_sln_02_01} & \includegraphics[scale=0.5]{gen_sln_02_02} \\
		$\gamma_{1}=1$                             & $\gamma_{1}=9$                             \\
		$\gamma_{2}=1$                             & $\gamma_{2}=-3$                             \\
		$\gamma_{3}=1$                            & $\gamma_{3}=-3$                             \\
	\end{tabular}
	\caption{Two parametric plots for the linearized system around $(c,-\frac{c}{a}, \frac{c}{a})$ with different initial values}
	\label{fig:nzero_sln}
\end{figure}

### 2016-10-30 padics/hw5.tex
working on rossler system

	\item Suppose that $k\in\NN$.
	      \begin{enumerate}
	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}\frac{1}{n^{k}}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.

	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}n^{k}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
	      \end{enumerate}
\end{enumerate}
\end{document}

>>>>

	\item Suppose that $k\in\NN$.
	      \begin{enumerate}
	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}\frac{1}{n^{k}}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
				  \begin{quote}
					  \textit{\textbf{Proof:}} Suppose otherwise. Then
					  \[
					  	\lim_{n\to\infty}\left|%
							\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}%
						\right|_{p} = 0.
					  \]
					  We can analyze the $p$-adic absolute value inside the limit:
					  \begin{align*}
						  \left|\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}\right|_{p} &\leq
						  \max\left\{%
						  	p^{-v_{p}\left(1/(n+1)^{k}\right)}, p^{-v_{p}\left(1/n^{k}\right)}%
						  \right\}
					  \end{align*}
					  We can solve for the valuation in the first term:
					  \[
					  	v_{p}\left(
							\frac{1}{(n+1)^{k}}
						\right) =
						-\frac{
							k\log{(n+1)}+\log{m}
						}{
							\log{p}
						},
					  \]
					  where $m\nmid p$. If we take the limit as $n\to\infty$ of this expression, we see that it tends towards negative infinity, so the $p$-adic absolute value tends towards infinity. Similarly, we can evaluate the limit of $|1/n^{k}|_{p}$, and we will see that it too tends towards infinity. Therefore,
					  \[
					  	\lim_{n\to\infty}\left|%
							\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}%
						\right|_{p} \neq 0,
					  \]
					  so the series does not converge since its sequence of partial sums does not converge.
				  \end{quote}

	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}n^{k}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
				  \begin{quote}
					  \textit{\textbf{Proof:}} We want to show that
					  \[
					  	\lim_{n\to\infty}\left|(n+1)^{k} - n^{k}\right|_{p}\neq 0.
					  \]
					  We can use a similar trick as above:
					  \begin{align*}
						  \left|(n+1)^{k} - n^{k}\right|_{p} &\leq \max\left\{

						  \right\}
					  \end{align*}
				  \end{quote}
	      \end{enumerate}
\end{enumerate}
\end{document}
